 

### 2021-05-06 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

~~In general, DGVMs range from Individual Based Models (IBMs) on one end of the complexity and scale spectrum to big-leaf models on the other end. Most climate models incorporate vegetation dynamics using big-leaf models, which include many first generation DGVMs that reduce 3D vegetation structure across the entire biosphere into a single vegetation layer. This simplification results in greater computational efficiency and thus easier integration into ESMs (Arora & Boer, 2010; Bonan et al. 2003; Cox, 2001; Krinner et al., 2005), although it comes at a cost of a lack of representation of important demographic processes, canopy gap formation, vertical light competition, competitive exclusion, and successional recovery from disturbance (Feeley et al., 2007; Hurtt et al. 1998; Moorcroft et al., 2001; Smith et al. 2001; Stark et al., 2012). On the other end of the spectrum are IBMs, also referred to as forest gap models, representing vegetation at the level of individual plants (Christoffersen et al., 2016; Fischer et al., 2016; Fyllas et al., 2014; Sato et al. 2007; Shuman et al. 2014; Smith et al., 2001; Bugmann, 2001; Dietze & Latimer, 2011). These models are often stochastic and spatially explicit, capturing spatial variability in the light environment and microclimates at high vertical and horizontal resolution. As a result, however, IBMs are computationally expensive and tend to be run at very local scales.~~

**The computationally feasible approach to representing vertical structuring in DVGMs lies in Cohort-based models (CBMs), which sit in the middle of this continuum between the oversimplified vegetation dynamics in big-leaf models and the computational expense of individual-based models.** CBMs represent vegetation as cohorts of individual plants, grouped together based on properties including size, age, and functional type (Haverd et al., 2013; Hurtt et al., 1998; Lischke et al. 2006; Medvigy et al. 2009; Moorcroft et al., 2001; Scherstjanoi et al. 2014; Smith et al., 2014; Weng et al., 2015). ~~Although this cohort approach serves as a compromise that incurs lower computational costs than IBMs while capturing many of the same dynamics, CBMs do not incorporate the stochastic processes of IBMs that typically allow for greater representation of functional diversity (Fisher et al., 2010).~~

*(KAT: I think all of the following content is great, but probably needs to be heavily tightened:)*

>>>>

**The computationally feasible approach to representing vertical structuring in DVGMs lies in Cohort-based models (CBMs), which sit in the middle of this continuum between the oversimplified vegetation dynamics in big-leaf models and the computational expense of individual-based models.** CBMs represent vegetation as cohorts of individual plants, grouped together based on properties including size, age, and functional type (Haverd et al., 2013; Hurtt et al., 1998; Lischke et al. 2006; Medvigy et al. 2009; Moorcroft et al., 2001; Scherstjanoi et al. 2014; Smith et al., 2014; Weng et al., 2015). 
*(KAT: I think all of the following content is great, but probably needs to be heavily tightened:)*
